Analysis of cloud enhancement events in a 30-year record of global solar irradiance at Thessaloniki, Greece

### 2024-09-12 article/article.Rmd
Commit 2024-09-12 13:11

```{r CLB-daily, echo=FALSE, fig.cap="Time-series of daily mean GHI measured at Thessaloniki for the period 1994 -- 2023."}
knitr::include_graphics("../images/P_daily_trend-5.png")
```

@@

<!--
```{r CS-change, echo=FALSE, fig.cap="Simulated long-term change in cloud-free irradiance relative to the climatological values due to changes in AOD in Thessaloniki for the period 1994 -- 2023."}
knitr::include_graphics("../images/P-CS-change-1.png")
```
-->

@@

```{r validation-GHI, echo=FALSE, fig.cap="Scatter plot of the measured GHI and the $\\text{GHI}_\\text{ref}$, for cloud-free days with data availability of more than $80\\,\\%$. The linear regression on the data is also shown on the figure with the blue line.", fig.pos='H'}
knitr::include_graphics("../images/P-validation-cloudfree-GHI-1.png")
```

@@

```{r example-day, echo=FALSE, fig.cap="Example of CE identification in Thessaloniki for 2019-07-11. The green line with blue symbols depicts the measured GHI in one-minute steps. The red line shows the modeled threshold for the detection of CE events, which are denoted with red circles. The black line represents the TOA TSI on a horizontal plane, equivalent to the threshold for the identification of ECE events, shown with magenta circles. The purple line is the modeled cloud-free irradiance. The dark yellow line is the solar constant of $1367\\,\\text{W}/\\text{m}^{2}$. The three sky images below the figure are taken at three characteristic instances (a, b, c) denoted on the figure with vertical orange dashed lines.", fig.pos='H'}
knitr::include_graphics("../images/P-example-day-3.png")
```

```{r example-year, echo=FALSE, fig.cap="Example scatter plot of the measured GHI and the reference cloud-free irradiance in Thessaloniki for the year 2005. The over-irradiance for CE and ECE events is color coded, while the remaining data points are shown in black. The reference green line goes through the origin with a slope of unity.", fig.pos='H'}
knitr::include_graphics("../images/P-example-years-12.png")
```

@@

```{r P-energy, echo=FALSE, fig.cap="Time-series of (a) the yearly CE number of occurrences, (b) the yearly mean OI and (c) the yearly excess irradiation at Thessaloniki for the period 1994 – 2023. The black lines represent the linear trends on the yearly data."}
knitr::include_graphics("../images/P-energy-complete-multi-3.png")
```

@@

```{r relative-month-occurrences, echo=FALSE, fig.cap="Seasonal variability of the relative occurrence of CE events in Thessaloniki for the period 1994 -- 2023, in the form of a box and whisker plot. The box contains the data between the lower $25\\,\\%$ and the upper $75\\,\\%$ percentiles, with the thick horizontal line representing the median. The vertical lines (whiskers) extend between the maximum and minimum monthly values."}
knitr::include_graphics("../images/P-CE-climatology-normlz-1.png")
```

<!--
```{r relative-month-occurrence, echo=FALSE, fig.cap="Seasonal variability of the number of CE events in Thessaloniki for the period 1994 -- 2023 normalized to the maximum occurring in June, in the form of a box and whisker plot. The monthly values have also been normalized to the relative abundance of valid GHI observations. The box contains the data between the lower $25\\,\\%$ and the upper $75\\,\\%$ percentiles, the thick horizontal line and the diamond symbol represent the median and the mean values, respectively. The vertical lines (whiskers) extend between the maximum and minimum monthly values."}
knitr::include_graphics("../images/clim-CE-month-norm-MAX-median-N-MW-1.png")
```
-->

@@

```{r ovir-distribution, echo=FALSE, fig.cap="Relative frequency distribution of CE events in Thessaloniki for the period 1994 -- 2023 as a function of OI. The histogram was split in two plots with different y-axis scales for better readability."}
knitr::include_graphics("../images/P-relative-distribution-diff-2.png")
```

@@

```{r ceg-duration-distribution, echo=FALSE, fig.cap="Relative frequency distribution of CE groups of consequent CE events according to their duration for Thessaloniki in the period 1994 -- 2023. The histogram was split in two plots with different y-axis scales for better readability."}
# knitr::include_graphics("../images/groups-1.png")
knitr::include_graphics("../images/groups-7.png")
```

@@

```{r group-2d, echo=FALSE, fig.cap="Relation of excess irradiation of CE groups with their duration for Thessaloniki in the period 1994 -- 2023. The logarithmic color scale denotes the frequency of the respective groups. Note that 19 pints with duration higher than 60 min are not shown."}
knitr::include_graphics("../images/P-groups-bin2d-1.png")
```

@@

```{r relative-month-occurancies-ECE, echo=FALSE, fig.cap="Seasonal variability of the relative occurrence of ECE events in Thessaloniki for the period 1994 -- 2023, in the form of a box and whisker plot. The box contains the data between the lower $25\\,\\%$ and the upper $75\\,\\%$ percentiles, with the thick horizontal line representing the median. The vertical lines (whiskers) extend between the maximum and minimum monthly values."}
knitr::include_graphics("../images/P-ECE-climatology-normlz-1.png")
```
<!--
```{r relative-month-occurancies-ECEs, echo=FALSE, fig.cap="Seasonal variability of the number of ECE events in Thessaloniki for the period 1994 -- 2023 normalized to the maximum occurring in March, in the form of a box and whisker plot. The box contains the data between the lower $25\\,\\%$ and the upper $75\\,\\%$ percentiles. The thick horizontal line and the diamond symbol represent the median and the mean values, respectively. The vertical lines (whiskers) extend between the maximum and minimum monthly values."}
knitr::include_graphics("../images/clim-ECE-month-norm-MAX-median-N-2.png")
```
-->

@@

```{r P-extreme-distribution, echo=FALSE, fig.cap="Relative frequency distribution of ECE events, in Thessaloniki for the period 1994 -- 2023, as a function of the extra irradiance over TSI at TOA for the same SZA."}
knitr::include_graphics("../images/P-extreme-distribution-2.png")
```

>>>>

```{r CLB-daily, echo=FALSE, fig.cap="Time-series of daily mean GHI measured at Thessaloniki for the period 1994 -- 2023."}
knitr::include_graphics("../images/P_daily_trend-5.pdf")
```

@@

<!--
```{r CS-change, echo=FALSE, fig.cap="Simulated long-term change in cloud-free irradiance relative to the climatological values due to changes in AOD in Thessaloniki for the period 1994 -- 2023."}
knitr::include_graphics("../images/P-CS-change-1.pdf")
```
-->

@@

```{r validation-GHI, echo=FALSE, fig.cap="Scatter plot of the measured GHI and the $\\text{GHI}_\\text{ref}$, for cloud-free days with data availability of more than $80\\,\\%$. The linear regression on the data is also shown on the figure with the blue line.", fig.pos='H'}
knitr::include_graphics("../images/P-validation-cloudfree-GHI-1.pdf")
```

@@

```{r example-day, echo=FALSE, fig.cap="Example of CE identification in Thessaloniki for 2019-07-11. The green line with blue symbols depicts the measured GHI in one-minute steps. The red line shows the modeled threshold for the detection of CE events, which are denoted with red circles. The black line represents the TOA TSI on a horizontal plane, equivalent to the threshold for the identification of ECE events, shown with magenta circles. The purple line is the modeled cloud-free irradiance. The dark yellow line is the solar constant of $1367\\,\\text{W}/\\text{m}^{2}$. The three sky images below the figure are taken at three characteristic instances (a, b, c) denoted on the figure with vertical orange dashed lines.", fig.pos='H'}
knitr::include_graphics("../images/P-example-day-3.pdf")
```

```{r example-year, echo=FALSE, fig.cap="Example scatter plot of the measured GHI and the reference cloud-free irradiance in Thessaloniki for the year 2005. The over-irradiance for CE and ECE events is color coded, while the remaining data points are shown in black. The reference green line goes through the origin with a slope of unity.", fig.pos='H'}
knitr::include_graphics("../images/P-example-years-12.pdf")
```

@@

```{r P-energy, echo=FALSE, fig.cap="Time-series of (a) the yearly CE number of occurrences, (b) the yearly mean OI and (c) the yearly excess irradiation at Thessaloniki for the period 1994 – 2023. The black lines represent the linear trends on the yearly data."}
knitr::include_graphics("../images/P-energy-complete-multi-3.pdf")
```

@@

```{r relative-month-occurrences, echo=FALSE, fig.cap="Seasonal variability of the relative occurrence of CE events in Thessaloniki for the period 1994 -- 2023, in the form of a box and whisker plot. The box contains the data between the lower $25\\,\\%$ and the upper $75\\,\\%$ percentiles, with the thick horizontal line representing the median. The vertical lines (whiskers) extend between the maximum and minimum monthly values."}
knitr::include_graphics("../images/P-CE-climatology-normlz-1.pdf")
```

<!--
```{r relative-month-occurrence, echo=FALSE, fig.cap="Seasonal variability of the number of CE events in Thessaloniki for the period 1994 -- 2023 normalized to the maximum occurring in June, in the form of a box and whisker plot. The monthly values have also been normalized to the relative abundance of valid GHI observations. The box contains the data between the lower $25\\,\\%$ and the upper $75\\,\\%$ percentiles, the thick horizontal line and the diamond symbol represent the median and the mean values, respectively. The vertical lines (whiskers) extend between the maximum and minimum monthly values."}
knitr::include_graphics("../images/clim-CE-month-norm-MAX-median-N-MW-1.pdf")
```
-->

@@

```{r ovir-distribution, echo=FALSE, fig.cap="Relative frequency distribution of CE events in Thessaloniki for the period 1994 -- 2023 as a function of OI. The histogram was split in two plots with different y-axis scales for better readability."}
knitr::include_graphics("../images/P-relative-distribution-diff-2.pdf")
```

@@

```{r ceg-duration-distribution, echo=FALSE, fig.cap="Relative frequency distribution of CE groups of consequent CE events according to their duration for Thessaloniki in the period 1994 -- 2023. The histogram was split in two plots with different y-axis scales for better readability."}
# knitr::include_graphics("../images/groups-1.pdf")
knitr::include_graphics("../images/groups-7.pdf")
```

@@

```{r group-2d, echo=FALSE, fig.cap="Relation of excess irradiation of CE groups with their duration for Thessaloniki in the period 1994 -- 2023. The logarithmic color scale denotes the frequency of the respective groups. Note that 19 pints with duration higher than 60 min are not shown."}
knitr::include_graphics("../images/P-groups-bin2d-1.pdf")
```

@@

```{r relative-month-occurancies-ECE, echo=FALSE, fig.cap="Seasonal variability of the relative occurrence of ECE events in Thessaloniki for the period 1994 -- 2023, in the form of a box and whisker plot. The box contains the data between the lower $25\\,\\%$ and the upper $75\\,\\%$ percentiles, with the thick horizontal line representing the median. The vertical lines (whiskers) extend between the maximum and minimum monthly values."}
knitr::include_graphics("../images/P-ECE-climatology-normlz-1.pdf")
```
<!--
```{r relative-month-occurancies-ECEs, echo=FALSE, fig.cap="Seasonal variability of the number of ECE events in Thessaloniki for the period 1994 -- 2023 normalized to the maximum occurring in March, in the form of a box and whisker plot. The box contains the data between the lower $25\\,\\%$ and the upper $75\\,\\%$ percentiles. The thick horizontal line and the diamond symbol represent the median and the mean values, respectively. The vertical lines (whiskers) extend between the maximum and minimum monthly values."}
knitr::include_graphics("../images/clim-ECE-month-norm-MAX-median-N-2.pdf")
```
-->

@@

```{r P-extreme-distribution, echo=FALSE, fig.cap="Relative frequency distribution of ECE events, in Thessaloniki for the period 1994 -- 2023, as a function of the extra irradiance over TSI at TOA for the same SZA."}
knitr::include_graphics("../images/P-extreme-distribution-2.pdf")
```
